---
title: "Introduction to scmisc"
author: "Diego Diez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to scmisc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(dplyr)
library(scmisc)

set.seed(1)
```


# Installation

You can install the release version of scmisc from [Github](https://github.com/ddiez/scmisc) with:

``` r
remotes::install_github("ddiez/scmisc")
```

# Toy dataset

This toy dataset contains 493 cells from mice in a _SingleCellExperiment_ object.

```{r}
sce
```


We can create some grouping variables.

```{r}
sce$group1 <- sample(c("A", "B"), ncol(sce), replace = TRUE)
sce$group2 <- sample(c("C", "D"), ncol(sce), replace = TRUE)
```


# Preprocess

To calculate PCA we can use `reduce_dim`.

```{r}
sce <- reduce_dim(sce, method = "pca")
```

# Plots

There are several plotting functions that work with _Seurat_ and _SingleCellExperiment_ objects.

## Scatter plot

```{r fig.width=6}
plot_coord(sce)
plot_coord(sce, color = "group1")
plot_coord(sce, color = "group2")
```

Expand by grouping variable.

```{r fig.width=6}
plot_coord(sce, expand = "group1")

plot_coord(sce, expand = c("group1", "group2"))
```


## Heatmap

The function `plot_heatmap` is a wrapper around `ComplexHeatmap::Heatmap` that simplifies plotting SingleCellExperiment and Seurat objects.

```{r fig.width=6, fig.height=4, message=FALSE}
plot_heatmap(sce[1:20, ])
```

We can use _ComplexHeatmap_ advance annotation features with the heatmap by passing them directly.

```{r fig.width=6, fig.height=4, message=FALSE}
cols <- circlize::colorRamp2(c(-2, 0, 2), c("purple", "black", "yellow"))
plot_heatmap(sce[1:20, ], col = cols)
```


```{r fig.width=6, fig.height=4, message=FALSE}
plot_heatmap(sce[1:20, ], column_split = sce$group1)
```

For column annotations we can directly pass the variable names to use with the color specification using _top_ann_ and _top_ann_col_.

```{r fig.width=6, fig.height=5, message=FALSE}
col <- list(group1 = c(A = "lightgrey", B = "black"),
            group2 = c("C" = "limegreen", "D" = "violetred"))
plot_heatmap(sce[1:20, ], column_split = sce$group1, top_ann = c("group1", "group2"), top_ann_col = col)
```

